Hydrodynamics and rheology of weakly adhesive vesicle doublets

### 2019-01-30 vesicle_code/docs/adhesion/adhesion.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in an extensional flow} 
\label{sec:eflow} 
The hydrodynamics of a single vesicle in an extensional flow has
revealed novel nonlinear vesicle dynamics not found for a viscous drop
\cite{KantslerSegreSteinberg2008_PRL, ZhaoShaqfeh2013_JFM,
Narsimhan2014_JFM,  DahlNarsimhanGouveia2016_SoftMatt}.  The stagnation
point (zero-velocity point) in an extensional flow, $\uu =\chi(-x,y) $,
is a saddle point where two laminar streams converge along an axis and
then diverge in a different direction.  With an active control algorithm
to fix the position of the stagnation point by adjusting the streaming
flow strength with a feedback loop~\cite{Johnson-Chavarria2011_EMJ}, a
particle can be trapped at the stagnation point for long time scales to
facilitate image acquisition or other detailed measures such as PIV of
flow inside and around the particle.  To explore the application of such
a fluid trap to measure the adhesion strength between two bound vesicle
membranes, we propose the following thought experiment.

We begin from two vesicles forming a doublet under an adhesion potential
in a quiescent flow as described in Section~\ref{sec:qflow}.  Once a
doublet forms and reaches an equilibrium configuration, we turn on the
fluid trap with the stagnation point placed at the center of the vesicle
doublet.  For simplification we assume that each vesicle in the doublet
has the same reduced area, membrane area, and bending stiffness.  At low
extensional rate, $\chi$, we expect the vesicle doublet to stay bound at
the fluid trap stagnation point.  On the other hand, the vesicle doublet
may become unstable and eventually separate at higher extensional rates.
Thus we expect there to exist a critical extensional rate above which
the vesicle doublet cannot stay bound under a given adhesion potential.
The dependence of the critical extensional rate on the adhesion
potential and the mechanical vesicle properties thus provides a means to
probe the physics of membrane adhesion.

We consider six pairs of vesicles with reduced areas 0.70, 0.75, 0.80,
0.85, 0.90, and 0.95, and we use extensional rates between $10^{-2}$ and
$10^{-1}$.  For all the simulations, the length of the vesicles are
fixed, while the area is changed to achieve the desired reduced area.
Since the stagnation point can be controlled in an experimental
setting~\cite{Johnson-Chavarria2011_EMJ}, we translate the vesicles at
each time step so that the stagnation point occurs exactly in the middle
of the two vesicles.  With this adjustment, the doublet can not migrate
towards $(0,\pm \infty)$; instead, for all time, each vesicle is
centered at $\pm(x_0,y_0)$ for some $(x_0,y_0)$.

@@

%In Figures~\ref{fig:extensional1}--\ref{fig:extensional3}, we show
%snapshots of two vesicles of reduced area $0.70$ in an extensional flow
%with three different extensional rates.  Interestingly, when the
%extensional rate does not overcome the adhesion rate, the doublet does
%not remain parallel to the stagnation point, but instead tilts to an
%inclination angle (Figures~\ref{fig:extensional1}
%and~\ref{fig:extensional2}.

\begin{figure}[htp]
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image01.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image02.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image03.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image04.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image05.png}
  \caption{\label{fig:extensional1} Two vesicles in an extensional flow.
  The reduced area is $0.7$, the Hamaker constant is $H = 0.7$, and the
  adhesion length scale is $\delta = 0.4$, extensional rate is $\chi =
  0.02$.  The center of the vesicle, denoted with a black square, is a
  stagnation point.}
\end{figure}

\begin{figure}[htp]
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image01.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image02.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image03.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image04.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image05.png}
  \caption{\label{fig:extensional2} Two vesicles in an extensional flow.
  The reduced area is $0.7$, the Hamaker constant is $H = 0.7$, and the
  adhesion length scale is $\delta = 0.4$, extensional rate is $\chi =
  0.07$.  The center of the vesicle, denoted with a black square, is a
  stagnation point.}
\end{figure}

\begin{figure}[htp]
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image01.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image02.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image03.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image04.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image05.png}
  \caption{\label{fig:extensional4} Two vesicles in an extensional flow.
  The reduced area is $0.7$, the Hamaker constant is $H = 0.7$, and the
  adhesion length scale is $\delta = 0.4$, extensional rate is $\chi =
  0.1$.  The center of the vesicle, denoted with a black square, is a
  stagnation point.}
\end{figure}

\begin{figure}[htp]
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image01.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image02.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image03.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image04.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image05.png}
  \caption{\label{fig:extensional3} Two vesicles in an extensional flow.
  The reduced area is $0.8$, the Hamaker constant is $H = 0.7$, and the
  adhesion length scale is $\delta = 0.4$, extensional rate is $\chi =
  0.1$.  Initially, the doublet tilts to an inclination angle, and then
  the extensional rate overcomes the adhesive force and the doublet is
  broken.  The center of the vesicle, denoted with a black square, is a
  stagnation point.}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in an extensional flow} 
\label{sec:eflow} 
The hydrodynamics of a single vesicle in an extensional flow has
revealed novel nonlinear vesicle dynamics not found for a viscous drop
\cite{KantslerSegreSteinberg2008_PRL, ZhaoShaqfeh2013_JFM,
Narsimhan2014_JFM, DahlNarsimhanGouveia2016_SoftMatt}.  The stagnation
point (zero-velocity point) in an extensional flow, $\uu =\chi(-x,y)$,
is a saddle point where two laminar streams converge along an axis and
then diverge in a different direction.  With an active control algorithm
to fix the position of the stagnation point by adjusting the streaming
flow strength with a feedback loop~\cite{Johnson-Chavarria2011_EMJ}, a
particle can be trapped at the stagnation point for long time scales to
facilitate image acquisition or other detailed measures such as PIV of
flow inside and around the particle.  To explore the application of such
a fluid trap to measure the adhesion strength between two bound vesicle
membranes, we propose the following thought experiment.

We begin from two vesicles forming a doublet under an adhesion potential
in a quiescent flow as described in Section~\ref{sec:qflow}.  Once a
doublet forms and reaches an equilibrium configuration, we turn on the
fluid trap with the stagnation point placed at the center of the vesicle
doublet.  For simplification we assume that each vesicle in the doublet
has the same reduced area, membrane area, and bending stiffness.  At low
extensional rate, $\chi$, we expect the vesicle doublet to stay fixed at
the fluid trap stagnation point.  On the other hand, the vesicle doublet
may become unstable and eventually separate at higher extensional rates.
Thus we expect there to exist a critical extensional rate above which
the vesicle doublet cannot stay bound under a given adhesion potential.
The dependence of the critical extensional rate on the adhesion
potential and the mechanical vesicle properties thus provides a means to
probe the physics of membrane adhesion.

\begin{figure}[htp]
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image01.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image02.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image03.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image04.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image05.png}
  \caption{\label{fig:extensional1} Two vesicles in an extensional flow.
  The reduced area is $A = 0.7$, the Hamaker constant is $H = 0.7$, and
  the adhesion length scale is $\delta = 0.4$, extensional rate is $\chi
  = 0.02$.  The center of the vesicle, denoted with a black square, is a
  stagnation point.}
\end{figure}

\begin{figure}[htp]
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image01.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image02.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image03.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image04.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image05.png}
  \caption{\label{fig:extensional2} Two vesicles in an extensional flow.
  The reduced area is $A = 0.7$, the Hamaker constant is $H = 0.7$, and
  the adhesion length scale is $\delta = 0.4$, extensional rate is $\chi
  = 0.07$.  The center of the vesicle, denoted with a black square, is a
  stagnation point.}
\end{figure}

\begin{figure}[htp]
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image01.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image02.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image03.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image04.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image05.png}
  \caption{\label{fig:extensional4} Two vesicles in an extensional flow.
  The reduced area is $A = 0.7$, the Hamaker constant is $H = 0.7$, and
  the adhesion length scale is $\delta = 0.4$, extensional rate is $\chi
  = 0.1$.  The center of the vesicle, denoted with a black square, is a
  stagnation point.}
\end{figure}

\begin{figure}[htp]
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image01.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image02.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image03.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image04.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image05.png}
  \caption{\label{fig:extensional3} Two vesicles in an extensional flow.
  The reduced area is $A = 0.8$, the Hamaker constant is $H = 0.7$, and
  the adhesion length scale is $\delta = 0.4$, extensional rate is $\chi
  = 0.1$.  Initially, the doublet tilts to an inclination angle, and
  then the extensional rate overcomes the adhesive force and the doublet
  is broken.  The center of the vesicle, denoted with a black square, is
  a stagnation point.}
\end{figure}

We consider six pairs of vesicles with reduced areas 0.70, 0.75, 0.80,
0.85, 0.90, and 0.95, and we vary the extensional rate, $\chi$, between
$10^{-2}$ and $10^{-1}$.  The length of each of the vesicles is fixed,
and the area is changed to achieve the desired reduced area.  Since the
stagnation point can be controlled in an experimental
setting~\cite{Johnson-Chavarria2011_EMJ}, we translate the vesicles at
each time step so that the stagnation point occurs exactly in the middle
of the two vesicles.  With this adjustment, the vesicles either remain
as a doublet in the fluid trap centered around the stagnation point, or
the doublet is broken and the vesicles diverge apart from one another.

In Figures~\ref{fig:extensional1}--\ref{fig:extensional3}, snapshots of
a fluid trap containing a vesicle doublet of reduced area 0.7 are
simulated at different extensional rates.  While this double remains
trapped at around the stagnation point, a doublet with a reduced area of
0.8 is broken for a sufficiently large extensional rate
(Figure~\ref{fig:extensional4}).  In the left plot of
Figure~\ref{fig:extensionalPhaseDiagram}, parameter values with a blue
mark result in a fluid trap while the parameter values with a red marks
result in the vesicles separating.

For extensional rates and reduced areas that result in a fluid trap, the
initial configuration of the vesicles have the diverging stream
direction is in the parallel to the vesicles' short axis.  Since the
stable configuration of two vesicles in an extensional flow has the
converging direction parallel to the vesicles' short axis, we expect the
vesicles to align themselves with the long axis parallel to the
diverging direction.  We observe in 
Figures~\ref{fig:extensional1}--\ref{fig:extensional3} that the vesicles
do in fact undergo a rotation, but not by a complete 90 degrees.  To
further analyze this rotation, we define the inclination angle of the
doublet as described in the right plot of
Figure~\ref{fig:extensionalPhaseDiagram}.

@@
